Demo arXiv template

### 2022-11-02 paper/paper.qmd
polish the toy example on constructing SPI in Sect 3

Consider a commonly used drought index: Standardised Precipitation Index (SPI) [@spi]. Its construction involves the following steps: 1) aggregate monthly summed precipitation into a time scale $k$, 2) fit the aggregated series into a distribution to get the density values, and 3) convert the density values to normal quantiles. Using the pipeline approach, the index will be constructed as:

```{r eval = FALSE, echo = TRUE}
DATA %>% 
  ts_aggregate(scale = 12, col = prcp) %>% 
  normalising(dist = "gamma", col = prcp, fit_method = "lmom") %>% 
  augment(col = "prcp_agg")
```

*The exact code needs to be further turned: 1) use verb: ts_aggreate -> aggregate_ts?, normalising -> normalise, 2) using `var` instead of `col` since it may be confused with "color", 3) normalising -> fit? since technically this step is no longer to normalise the fitted density value into normal quantiles but a distribution fitting step.*

Here the column `prcp` is first temporally *aggregated* with a time scale of 12. The output is then normalised with a gamma distribution, fitted using the L-moment. The `augment` step, following the verb in the `broom` package, accepts the fitted distributions and adds information about each month in the dataset to get the final index.

@@

compare <- res %>% left_join(from_SPI, by = "ym") 
compare %>% 
  ggplot(aes(x = SPI_idx, y = pipeline_idx)) + 
  geom_point() + 
  theme_bw() + 
  theme(legend.position = "bottom", 
        aspect.ratio = 1) +
  ggtitle("Index values of Standardised Precipitation Index (SPI)", 
          subtitle = "using wichita data from the SPEI package") + 
  xlab("Calculated from a single command") + 
  ylab("Calculated from pipeline")

```

```{r fig-toy-example, eval = TRUE, echo = FALSE}
#| fig-align: center
#| fig-cap: "this is the caption"
#| out-height: 50%
#| out-width: 50%
knitr::include_graphics(here::here("figures/toy-example-spi.png"))
```

While the calculated index value from the pipeline approach is the same as those calculated using a single line command like `spi(DATA, ARGUMENTS)`, as shown in @fig-toy-example, the pipeline approach gives more flexibility to perform diagnostics on the intermediate steps and to extend on the initial index calculation. For example, model goodness of fit test can be checked after `normalising()` through:

```{r eval = FALSE, echo = TRUE}
normalising(dist = "gamma", col = d, fit_method = "lmom") %>% 
  gof_test()
```

and bootstrap uncertainty on the index, with 100 boostrap samples, can be calculated using `n_boot = 100` in the `normalising()` as:

```{r eval = FALSE, echo = TRUE}
normalising(dist = "gamma", col = d, fit_method = "lmom", n_boot = 100) 
```

>>>>

Consider a commonly used drought index: Standardised Precipitation Index (SPI) [@spi]. Its construction involves: 1) aggregating monthly precipitation (`prcp`) with a time scale, 2) fitting the aggregated series with a chosen distribution to get density values, and 3) converting the density values to normal quantiles. Under the pipeline approach, SPI will be constructed as:

```{r eval = FALSE, echo = TRUE}
DATA %>% 
  aggregate(scale = 12, var = prcp) %>% 
  normalise(dist = gamma, var = prcp, fit_method = lmom) %>% 
  augment(var = .agg_prcp)
```

<!-- *FWI:*  -->

<!--   - *verbs as name for each step: `aggregate`, `normalise`* -->
<!--   - *use NSE for variable name: `prcp`, `c(tavg, prcp)`* -->
<!--   - *created variables are prefixed with dot: `.agg_prcp`* -->
<!--   - *avoid string and use function to avoid "gamma" vs. "gam": `gamma()`, `lmom()`* -->

Here each command corresponds to one step described above with additional arguments specific to the step. 
The pipeline construction produces the same index value as command like `spi(...)`, as shown in @fig-toy-example, but with additional benefit:

  - Multiple scales and multiple distributions can be fitted using the `c(...)` syntax to compare index values constructed from different parameterisations; 
  - Intermediate results can be checked after each step; and
  - Additional steps and analysis can be wired into the workflow for index diagnostics and customised user need.

@@

compare <- res %>% left_join(from_SPI, by = "ym") 
compare %>% 
  ggplot(aes(x = SPI_idx, y = pipeline_idx)) + 
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") + 
  geom_point() + 
  theme_bw() + 
  theme(legend.position = "bottom", 
        aspect.ratio = 1) +
  xlab("Index values from `spi()` in package `SPEI`") + 
  ylab("Index values from pipeline") + 
  theme(panel.grid.minor = element_blank())
```

```{r fig-toy-example, eval = TRUE, echo = FALSE}
#| fig-align: center
#| fig-cap: "Standardised Precipitation Index (SPI) calculated using two approaches: the pipeline construction (y-axis) and the function `spi()` from the `SPEI` package (x-axis), using the `wichita` data from the package `SPEI`. All the points align diagnally, indicating the same numerical result from both approaches."
#| out-height: 70%
#| out-width: 70%
knitr::include_graphics(here::here("figures/toy-example-spi.png"))
```

### 2022-11-02 paper/paper.tex
polish the toy example on constructing SPI in Sect 3

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, frame hidden, boxrule=0pt, sharp corners, enhanced, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

@@

Consider a commonly used drought index: Standardised Precipitation Index
(SPI) (McKee, Doesken, and Kleist 1993). Its construction involves the
following steps: 1) aggregate monthly summed precipitation into a time
scale \(k\), 2) fit the aggregated series into a distribution to get the
density values, and 3) convert the density values to normal quantiles.
Using the pipeline approach, the index will be constructed as:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{DATA }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{ts\_aggregate}\NormalTok{(}\AttributeTok{scale =} \DecValTok{12}\NormalTok{, }\AttributeTok{col =}\NormalTok{ prcp) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{normalising}\NormalTok{(}\AttributeTok{dist =} \StringTok{"gamma"}\NormalTok{, }\AttributeTok{col =}\NormalTok{ prcp, }\AttributeTok{fit\_method =} \StringTok{"lmom"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{augment}\NormalTok{(}\AttributeTok{col =} \StringTok{"prcp\_agg"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\emph{The exact code needs to be further turned: 1) use verb:
ts\_aggreate -\textgreater{} aggregate\_ts?, normalising -\textgreater{}
normalise, 2) using \texttt{var} instead of \texttt{col} since it may be
confused with ``color'', 3) normalising -\textgreater{} fit? since
technically this step is no longer to normalise the fitted density value
into normal quantiles but a distribution fitting step.}

Here the column \texttt{prcp} is first temporally \emph{aggregated} with
a time scale of 12. The output is then normalised with a gamma
distribution, fitted using the L-moment. The \texttt{augment} step,
following the verb in the \texttt{broom} package, accepts the fitted
distributions and adds information about each month in the dataset to
get the final index.

@@

{\centering \includegraphics[width=0.5\textwidth,height=0.5\textheight]{../figures/toy-example-spi.png}

@@

\caption{\label{fig-toy-example}this is the caption}

@@

While the calculated index value from the pipeline approach is the same
as those calculated using a single line command like
\texttt{spi(DATA,\ ARGUMENTS)}, as shown in
Figure~\ref{fig-toy-example}, the pipeline approach gives more
flexibility to perform diagnostics on the intermediate steps and to
extend on the initial index calculation. For example, model goodness of
fit test can be checked after \texttt{normalising()} through:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{normalising}\NormalTok{(}\AttributeTok{dist =} \StringTok{"gamma"}\NormalTok{, }\AttributeTok{col =}\NormalTok{ d, }\AttributeTok{fit\_method =} \StringTok{"lmom"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{gof\_test}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

and bootstrap uncertainty on the index, with 100 boostrap samples, can
be calculated using \texttt{n\_boot\ =\ 100} in the
\texttt{normalising()} as:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{normalising}\NormalTok{(}\AttributeTok{dist =} \StringTok{"gamma"}\NormalTok{, }\AttributeTok{col =}\NormalTok{ d, }\AttributeTok{fit\_method =} \StringTok{"lmom"}\NormalTok{, }\AttributeTok{n\_boot =} \DecValTok{100}\NormalTok{) }
\end{Highlighting}
\end{Shaded}

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

Consider a commonly used drought index: Standardised Precipitation Index
(SPI) (McKee, Doesken, and Kleist 1993). Its construction involves: 1)
aggregating monthly precipitation (\texttt{prcp}) with a time scale, 2)
fitting the aggregated series with a chosen distribution to get density
values, and 3) converting the density values to normal quantiles. Under
the pipeline approach, SPI will be constructed as:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{DATA }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{aggregate}\NormalTok{(}\AttributeTok{scale =} \DecValTok{12}\NormalTok{, }\AttributeTok{var =}\NormalTok{ prcp) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{normalise}\NormalTok{(}\AttributeTok{dist =}\NormalTok{ gamma, }\AttributeTok{var =}\NormalTok{ prcp, }\AttributeTok{fit\_method =}\NormalTok{ lmom) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{augment}\NormalTok{(}\AttributeTok{var =}\NormalTok{ .agg\_prcp)}
\end{Highlighting}
\end{Shaded}

Here each command corresponds to one step described above with
additional arguments specific to the step. The pipeline construction
produces the same index value as command like \texttt{spi(...)}, as
shown in Figure~\ref{fig-toy-example}, but with additional benefit:

\begin{itemize}
\tightlist
\item
  Multiple scales and multiple distributions can be fitted using the
  \texttt{c(...)} syntax to compare index values constructed from
  different parameterisations;
\item
  Intermediate results can be checked after each step; and
\item
  Additional steps and analysis can be wired into the workflow for index
  diagnostics and customised user need.
\end{itemize}

@@

{\centering \includegraphics[width=0.7\textwidth,height=0.7\textheight]{../figures/toy-example-spi.png}

@@

\caption{\label{fig-toy-example}Standardised Precipitation Index (SPI)
calculated using two approaches: the pipeline construction (y-axis) and
the function \texttt{spi()} from the \texttt{SPEI} package (x-axis),
using the \texttt{wichita} data from the package \texttt{SPEI}. All the
points align diagnally, indicating the same numerical result from both
approaches.}

@@
